\chapter{Amount Hiding}
\label{chapter:pedersen-commitments}

In most cryptocurrencies like Bitcoin, transaction output notes, which give spending rights to `amounts' of money, communicate those amounts in clear text. This allows observers to easily verify the amount spent equals the amount sent.

In MobileCoin we use {\em commitments} to hide output amounts from everyone except senders and receivers, while still giving observers confidence that a transaction sends no more or less than what is spent. As we will see, amount commitments must also have corresponding `range proofs' that prove the hidden amount is within a legitimate range.



\section{Commitments}
\label{sec:commitments}

Generally speaking, a cryptographic {\em commitment scheme} is a way of committing to a value without revealing the value itself. After committing to something, you are stuck with it.

For example, in a coin-flipping game Alice could privately commit to one outcome (i.e.\ ‘call it’) by hashing her committed value alongside secret data and publishing the hash. After Bob flips the coin, Alice declares which value she committed to and proves it by revealing the secret data. Bob could then verify her claim.

In other words, assume Alice has a secret string ``My secret" and the value she wants to commit to is $heads$. She hashes $h = \mathcal{H}(\textrm{``My secret"}, heads)$ and gives $h$ to Bob. Bob flips a coin, then Alice tells Bob the secret string ``My secret" and that she committed to $heads$. Bob calculates $h’ = \mathcal{H}(\textrm{``My secret"}, heads)$. If $h' = h$, then he knows Alice called $heads$ before the coin flip.

Alice uses the so-called `salt', ``My secret", so Bob can't just guess $\mathcal{H}(heads)$ and $\mathcal{H}(tails)$ before his coin flip, and figure out she committed to $heads$.\footnote{If the committed value is very difficult to guess and check, e.g.\ if it's an apparently random elliptic curve point, then salting the commitment isn't necessary.}



\section{Pedersen commitments}
\label{sec:pedersen-intro}

A {\em Pedersen commitment} \cite{Pedersen1992} is a commitment that has the property of being {\em additively homomorphic}. If \(C(a)\) and \(C(b)\) denote the commitments for values \(a\) and \(b\) respectively, then \(C(a + b) = C(a) + C(b)\).\footnote{Additively homomorphic in this context means addition is preserved when you transform scalars into EC points by applying, for scalar $x$, $x \rightarrow x G$. In other words, after transforming a scalar into a curve point, addition operations between points proceed in `parallel' to scalar additions. The transformation of scalar $a + b$ always equals $a G + b G$ based on the individual transformations of scalars $a$ and $b$.} This property will be useful when committing transaction amounts, as one can prove, for instance, that inputs equal outputs, without revealing the amounts at hand.
\\

Fortunately, Pedersen commitments are easy to implement with elliptic curve cryptography, as the following holds trivially \[a G + b G = (a + b) G\]

Clearly, by defining a commitment as simply \(C(a) = a G\), we could easily create cheat tables of commitments to help us recognize common values of $a$.

To attain information-theoretic privacy, one needs to add a secret {\em blinding factor} and another generator \(H\), such that it is unknown for which value of \(\gamma\) the following holds: \(H = \gamma G\). The hardness of the discrete logarithm problem ensures calculating $\gamma$ from $H$ is infeasible.\footnote{In the case of MobileCoin\marginnote{[MC-tx] src/ring\_sig- nature/mod.rs {\tt GENERATORS}}, $H = H_p(G)$.}

We can then define the commitment to a value \(a\) as \(C(x, a) = x G + a H\), where \(x\) is the blinding factor (a.k.a.\ `mask') that prevents observers from guessing $a$.

Commitment $C(x, a)$ is information-theoretically private because there are many possible combinations of $x$ and $a$ that would output the same $C$.\footnote{Basically, there are many $x’$ and $a’$ such that $x’+a’ \gamma = x+a \gamma$. A committer knows one combination, but an attacker has no way to know which one. This property is also known as `perfect hiding' \cite{adam-zero-to-bulletproofs}. Even the committer can't find another combination without solving the DLP for $\gamma$, a property called `computational binding' \cite{adam-zero-to-bulletproofs}.} If $x$ is truly random, an attacker would have literally no way to figure out $a$ \cite{maxwell-ct-2, SCOZZAFAVA1993313}.%{https://people.xiph.org/~greg/confidential_values.txt}



\section{Amount commitments}
\label{sec:pedersen-ringct}

In MobileCoin, output amounts are stored in transactions as Pedersen commitments. We define a commitment to an output’s amount $b$ as:\vspace{.175cm}
\[C(y,b) = y G + b H\marginnote{[dalekBP] src/gener- ators.rs {\tt Pedersen- Gens::com- mit()}}\]

Recipients should be able to know how much money is in each output they own, as well as reconstruct the amount commitments, so they can be used as the inputs to new transactions. This means the blinding factor $y$ and amount $b$ must be communicated to the receiver.

The solution adopted is a Diffie-Hellman shared secret $r_t K_B^{v,t}$ using the `transaction output public key' (recall Section \ref{sec:multi_out_transactions}). Every output in the blockchain has a mask $y_t$ that senders and receivers can privately compute, and a $\mathit{masked\_value}_t$ stored in the transaction's data. While $y_t$ is an elliptic curve scalar and occupies 32 bytes, $b_t$ will be restricted to 8 bytes by the range proof so only an 8-byte value needs to be stored.\footnote{Domain separation tags are written explicitly here to emphasize the hash results are not the same.}\footnote{We include the index $t$ (from within the transaction that created a given output) for clarity, but this information is lost when an output is published in a block, since outputs are sorted\marginnote{*consensus/ enclave/impl/ src/lib.rs {\tt form\_block()}} by txout public key within each block to ensure all validator nodes create the same block.}\marginnote{[MC-tx] src/amount/ mod.rs {\tt Amount:: new()}}[.725cm]\vspace{.175cm}
\begin{align*}
  y_t &= \mathcal{H}_n(``mc\_amount\_blinding", [r_t K_B^{v,i}]) \\
  \mathit{masked\_value}_t &= b_t \oplus_8 \mathcal{H}_n(``mc\_amount\_value”, [r_t K_B^{v,i}])
\end{align*}

Here, $\oplus_8$ means to perform an XOR operation (Section \ref{sec:XOR_section}) between the first 8 bytes of each operand ($b_t$ which is already 8 bytes, and $\mathcal{H}_n(...)$ which is 32 bytes). Recipients can perform the same XOR operation on $\mathit{masked\_value}_t$ to reveal $b_t$.

The receiver Bob will be able to calculate\marginnote{[MC-tx] src/amount/ mod.rs {\tt Amount:: get\_value()}} the blinding factor $y_t$ and the amount $b_t$ using the txout public key $r_t K^{s,i}$ and his view key $k_B^v$. He can also check that the commitment $C(y_t, b_t)$ provided in the output data, henceforth denoted $C_t^b$, corresponds to the amount at hand.\\

More generally, any third party with access to Bob’s view key could decrypt his output amounts, and also make sure they agree with their associated commitments.



\section{RingCT introduction}
\label{sec:ringct-introduction}

A transaction will contain copies of other transactions' outputs (telling validators which old outputs are to be spent), and its own outputs. The content of an output includes a one-time address (assigning ownership of the output), a txout public key (for accessing the output via Diffie-Hellman exchange), an output commitment hiding the amount, the encoded output amount from Section \ref{sec:pedersen-ringct}, and a so-called `encrypted fog hint' (an 84-byte field we discuss in Section \ref{subsec:fog-encrypted-fog-hints}).

While a transaction's verifiers don’t know how much money is contained in each input and output, they still need to be sure the sum of input amounts equals the sum of output amounts. MobileCoin uses a technique called RingCT \cite{MRL-0005-ringct} to accomplish this.

If we have a transaction with $m$ inputs containing amounts \(a_1, ..., a_m\), and $p$ outputs with amounts \(b_1, ..., b_p\), then an observer would justifiably expect that:\footnote{If the intended total output amount doesn't precisely equal any combination of owned outputs, then transaction authors can add a `change' output sending extra money back to themselves. By analogy to cash, with a 20\$ bill and 15\$ expense you will receive 5\$ back from the cashier.}\vspace{.175cm}
\[\sum_j a_j - \sum_t b_t = 0\]

Since commitments are additive and we don't know $\gamma$, we could easily prove our inputs equal outputs to observers by making the sum of commitments to input and output amounts equal zero (i.e.\ by setting the sum of output blinding factors equal to the sum of old input blinding factors):\footnote{Recall from Section \ref{subsec:elliptic_curves_section} we can subtract a point by inverting its coordinates then adding it. If $P = (x, y)$, $-P = (-x, y)$. Recall also that negations of field elements are calculated$\pmod q$, so $(–x \pmod q)$.}\vspace{.175cm}
\[\sum_{j}{C_{j, in}} - \sum_{t}{C_{t, out}} = 0\]

To avoid sender identifiability we use a slightly different approach. The amounts being spent correspond to the outputs of previous transactions, which had commitments\vspace{.175cm}
\[C^a_{j} = x_j G + a_j H\]

The sender can create new commitments to the same amounts but using different blinding factors; that is,
\[C'^a_{j} = x'_j G + a_j H\]

Clearly, she would know the private key of the difference between the two commitments: \vspace{.175cm}
\[C^a_{j} - C'^a_{j} = (x_j - x'_j) G\]\\
Hence, she would be able to use this value as a {\em commitment to zero}, since she can make a signature with the private key $(x_j - x'_j) = z_j$ and prove there is no $H$ component to the sum (assuming $\gamma$ is unknown). In other words prove that $C^a_{j} - C'^a_{j} = z_j G + 0H$, which we will actually do in Chapter~\ref{chapter:transactions} when we discuss the structure of RingCT transactions.

Let us call $C'^a_j$ a {\em pseudo output commitment}. Pseudo output commitments are included in transaction data, and there is one for each input.

Before committing the outputs in a transaction to the blockchain, the network will want to verify that amounts balance. Blinding factors for pseudo and output commitments are selected such that\vspace{.175cm}
\[\sum_j x'_j  - \sum_t y_t = 0\]

This\marginnote{[MC-tx] src/ring\_ signature/ rct\_bullet- proofs.rs {\tt verify()}} allows us to prove input amounts equal output amounts:\vspace{.175cm}
\[(\sum_j C'^a_{j} - \sum_t C^b_{t}) = 0\]

Fortunately, choosing such blinding factors is easy. In the current version of Mobilecoin, all blinding factors are random except for the $m$\nth pseudo out commitment, where $x'_m$ is simply
\[x'_m = \sum_t y_t - \sum_{j=1}^{m-1} x'_j\marginnote{{\tt sign\_with\_ balance\_ check()}}\]



\section{Range proofs}
\label{sec:range_proofs}

One problem with additive commitments is that, if we have commitments $C(a_1)$, $C(a_2)$, $C(b_1)$, and $C(b_2)$ and we intend to use them to prove that $(a_1 + a_2) - (b_1 + b_2) = 0$, then those commitments would still apply if one value in the equation were `negative'.

For instance, we could have $a_1 = 6$, $a_2 = 5$, $b_1 = 21$, and $b_2 = -10$.\vspace{.175cm}
\begin{flalign*}
    && (6 + 5) - (&21 + -10) = 0&\\
     \intertext{\quad \quad \quad \quad \quad where} && 21G + -10G = 21G + (&l-10)G = (l + 11)G = 11G&
\end{flalign*}

Since $-10 = l-10$, we have effectively created $l$ more MOB (over 7.2x10$^{74}$!) than we put in.

%Bulletproofs start here
The solution addressing this issue in MobileCoin is to prove each output amount is in a certain range (from 0 to $2^{64}-1$) using the Bulletproofs proving scheme first described by Benedikt B\"{u}nz {\em et al.} in \cite{Bulletproofs_paper} (and also explored in \cite{adam-zero-to-bulletproofs,dalek-bulletproofs-notes}).\footnote{It's conceivable that with several outputs in a legitimate range, the sum of their amounts could roll over and cause a similar problem. However, when the maximum output amount is much smaller than $l$ it takes a huge number of outputs for that to happen. For example, if the range is 0-5 and $l = 99$, then to counterfeit money using an input of 2, we would need $5 + 5 + …. + 5 + 1 = 101 \equiv 2 \pmod{99}$, for 21 total outputs. In Ed25519 $l$ is about 2\^{}189 times bigger than the available range, which means a ridiculous 2\^{}189 outputs to counterfeit money.} Given the involved and intricate nature of Bulletproofs, the scheme is not elucidated in this document. Moreover we feel the cited materials adequately present its concepts.%\footnote{When RingCT was invented \cite{MRL-0005-ringct}, range proofs were done with Borromean ring signatures \cite{Signatures2015BorromeanRS}, which are elaborated in Section 3.4 of \cite{ztm-1}.}

The Bulletproof proving algorithm\marginnote{[dalekBP] src/range\_ proof/mod.rs {\tt prove\_mult- iple\_with\_ rng()}} takes as input output amounts $b_t$ and commitment masks $y_t$, and outputs all $C^b_t$ and an $n$-tuple aggregate proof\footnote{Vectors $\mathbb{L}$ and $\mathbb{R}$ contain $\lceil \textrm{log}_2(64 \cdot p) \rceil$ elements each. $\lceil$ $\rceil$ means the log function is rounded up. Due to their construction, some Bulletproofs use `dummy outputs' as padding to ensure $p$ plus the number of dummy outputs is a power of 2. Those dummy outputs can be generated during verification\marginnote{[MC-tx] src/range\_ proofs/mod.rs {\tt check\_range\_ proofs()}}, and are not stored with the proof data.}\footnote{The variables in a Bulletproof are unrelated to other variables in this document. Symbol overlap is merely coincidental. Note that in MobileCoin transactions Bulletproofs are represented by a byte blob that gets passed to the {\tt dalek-cryptography bulletproofs} library \cite{dalek-bulletproofs-lib} for verification.}\vspace{.155cm} 
\[\Pi_{BP} = (A, S, T_1, T_2, t_x, t^{blinding}_x, e^{blinding}, \mathbb{L}, \mathbb{R}, a, b)\]

That single proof is used to prove all output amounts are in range at the same time, as aggregating them greatly reduces space requirements (although it does increase the time to verify).\footnote{It turns out multiple separate Bulletproofs can be `batched' together, which means they are verified simultaneously. Doing so improves how long it takes to verify them, and there is no theoretical limit to how many can be batched together. Currently in MobileCoin each transaction is only allowed one Bulletproof, which is validated by itself without batching\marginnote{[MC-tx] src/ring\_ signature/ rct\_bullet- proofs.rs {\tt verify()}}. As we will discuss more in the coming chapters, MobileCoin nodes discard Bulletproofs after validating them, so batching would only be beneficial if transaction volume is significant.} The verification algorithm takes as input all $C^b_t$, and $\Pi_{BP}$, and outputs {\tt true} if all committed amounts are in the range 0 to $2^{64} - 1$.

The $n$-tuple $\Pi_{BP}$ occupies $(2 \cdot \lceil \textrm{log}_2(64 \cdot (m + p)) \rceil + 9) \cdot 32$ bytes of storage.\footnote{\label{footnote:commitments-pseudo-output-commit-range-proofs}In the initial version of MobileCoin pseudo output commitments are also range proofed in the same way as output commitments, which is why there is an `$m$' in $(64 \cdot (m + p))$. Apparently the MobileCoin development team plans to stop range proofing pseudo output commitments in the next version of the protocol \cite{mc-pull-request-revert-remove-pseudo-output-from-range-proof}, which seems to us like a good idea since range proofing them serves no purpose.}